### 2008-03-17 SAiP/Module1/H3/doc/architecturaldesign.tex
Last corrections.

%%%
% Architectural Design
%%%
\section{Architectural Design -- Beer Web Store}
Consider the following main use case for a Beer Web Store for speciality
beers:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. The software architect and stakeholders of the system
have based on this outlined the following two central Quality
Attribute Scenarios:
\begin{itemize}
  \item A customer makes a payment to the system in normal mode. The
  payment is processed within 10 seconds
  \item The payment service fails during normal operation. The system
  detects this. The administrator is notified and the system continues
  in degraded mode until the payment service is made available
\end{itemize}

@@

When a system is Strictly Layered it gives rise to high \emph{Portability} (a for of \emph{Modifiability}), as it is possible to abstract all dependencies of a given layer by simply abstracting the layer directly beneath it, and if anything in a layer changes then only the layer directly above it must be updated.

This also enables simpler \emph{Testability}, as the introduction of test doubles (fake objects, stubs, spys and/or mocks) must only be performed on the in a single layer; the one directly below the one under test.

@@

\begin{itemize}
  \item Process or communicating processes.
	\begin{itemize}
	  \item This structure shows the execution path through a system, and is used to give the engineer an understand about the dynamic behaviour of the system. These structures may help engineers increase Performance and Availability.
	\end{itemize}
  \item Concurrency.
	\begin{itemize}
	  \item This structure is used to see the parallelism in the system, or part of the system, and when a system, or part of a system, is build around this structure it promotes performance. 
	\end{itemize}
  \item Shared data or Repository
	\begin{itemize}
	  \item This is a very good structure to use to illustrate how the data is processed and consumed to/from the database (Server to/from Database Fa\c ade). When a system, or part of a system, is build around this structure it promotes performance and data integrity. 
	\end{itemize}
  \item Client-server
	\begin{itemize}
  		\item This is a very good structure to see the protocols used between the Clients and the Server, and the messages they communicate. When a system, or part of a system, is build around this structure it promotes separation of concern (i.e. Modifiability) and allows for Availability tactics like Load balancing.
	\end{itemize}
\end{itemize}

In the Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Fa\c ade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).

@@

\fixme{Possible extension - insert diagrams}

>>>>

%%%
% Architectural Design
%%%
\section{Architectural Design -- Beer Web Store}
Consider the following main use case for a Beer Web Store for specialty
beers:
\begin{quote}
  The customer browses the beer catalogue for descriptions and
  reviews. The customer selects appropriate beers for purchase and
  these are added to his shopping basket. When the customer checks
  out, payment is validated through the payment service
\end{quote}
The main architectural drivers are considered to be performance and
availability. The software architect and stakeholders of the system
have based on this outlined the following two central Quality
Attribute Scenarios:
\begin{itemize}
  \item A customer makes a payment to the system in normal mode. The
  payment is processed within 10 seconds
  \item The payment service fails during normal operation. The system
  detects this. The administrator is notified and the system continues
  in degraded mode until the payment service is made available
\end{itemize}

@@

When a system is Strictly Layered it gives rise to high \emph{Portability} (a form of \emph{Modifiability}), as it is possible to abstract all dependencies of a given layer by simply abstracting the layer directly beneath it, and if anything in a layer changes then only the layer directly above it must be updated.

This also enables simpler \emph{Testability}, as the introduction of test doubles (fake objects, stubs, spys and/or mocks) must only be performed in a single layer; the one directly below the layer under test.

@@

\begin{itemize}
  \item Process or communicating processes.
	\begin{itemize}
	  \item This structure shows the execution path through a system, and is used to give the engineer an understanding about the dynamic behavior of the system. These structures may help engineers increase Performance and Availability.
	\end{itemize}
  \item Concurrency.
	\begin{itemize}
	  \item This structure is used to see the parallelism in the system, or part of the system, and when a system, or part of a system, is build around this structure it promotes performance. 
	\end{itemize}
  \item Shared data or Repository
	\begin{itemize}
	  \item This is a very good structure to use to illustrate how the data is processed and consumed to/from the database (Server to/from Database Fa\c cade). When a system, or part of a system, is build around this structure it promotes performance and data integrity. 
	\end{itemize}
  \item Client-server
	\begin{itemize}
  		\item This is a very good structure to see the protocols used between the Clients and the Server, and the messages they communicate. When a system, or part of a system, is build around this structure it promotes separation of concern (i.e. Modifiability) and allows for Availability tactics like Load balancing.
	\end{itemize}
\end{itemize}

The Beer Web Store has two clear C\&C structures, the Shared data structure between the Server and the Database Fa\c cade and the Client-server between the Client and Server (see Figure \ref{fig:module_view}).

@@

### 2008-03-17 SAiP/Module1/H3/doc/productlines.tex
Last corrections.

The above two approaches are defined as Heavyweight and Lightweight, respectively. A statistical analysis of these two approaches, compared to simply building the products from scratch, can be shown in Figure \ref{fig:product-line}.

\begin{figure}[!htb]
\centerline{\epsfig{figure=figures/product-line,scale=0.9 }}
\caption{Heavy-weight vs. Light-weight}
\label{fig:product-line}
\end{figure}

@@

Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well know that the architecture can be fully specified in advance.

@@

It may now seem like SOA is the answer to creating a product line, yet this is not the case. It is very possible to create very specific services, like the ``Beer look-up service'', and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line. The same architectural consideration must be done when creating the elements of a product line and define their interrelations, must also be done when defining the services and their interrelations. The only ``structure'' that SOA imposes by definition is the loose coupling between the user and the service.

>>>>

The above two approaches are defined as Heavyweight and Lightweight, respectively. A statistical analysis of these two approaches, compared to simply building the products from scratch, is shown in Figure \ref{fig:product-line}.

\begin{figure}[!htb]
\centerline{\epsfig{figure=figures/product-line,scale=0.9 }}
\caption{Heavy-weight vs. Light-weight \cite{mcgregor2002}}
\label{fig:product-line}
\end{figure}

@@

Based on this the Beer Web Store architect must have decided that the products in the pipe-line and the Beer Web Store share a common architecture and that the future products are so well known that the architecture can be fully specified in advance.

@@

It may now seem like SOA is the answer to creating a product line, yet this is not the case. It is very possible to create very specific services, like the ``Beer look-up service'', and this may be just fine for a Beer Web Store based on SOA, yet this is not a product line. The same architectural consideration must be done when creating the elements of a product line and define their interrelations, as when defining the services and their interrelations. The only ``structure'' that SOA imposes by definition is the loose coupling between the user and the service.

### 2008-03-17 SAiP/Module1/H3/doc/qualityattributes.tex
Adding label
Last corrections.

Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve. For examples sake lets asume that the above scenario was selected as top priority.

As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-resonsiveness and how many simultanious users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.

@@

Finding out if a resource is down may be done using one of the fault detection tactice; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not benificial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintance process on the same or another machine, and if one (or perhaps more) heartbeats is missed then the maintanance service can act.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and tehn if an error occur (as detected by the heartbeat) then restart or swap to another server/database in 30 seconds. As 30 seconds is insufficient to bott a webserver and a database, the Spare tactic would be insufficient. At the same tiem having Active redundancy would be over-kill, as a this kind of immidiat switch-over is not required. For these reasons Passive Redundancy is sufficient.

As it is an internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience beeing thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. for this it will be benificial to use the \emph{Rollback} tactic on the \emph{Basket} component.

Finally a web-server may be deployed multiple times on teh same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic. This is e.g. done by the Apache-server.

As it is not only the web server which may become non-responsive, it is also the database, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the ping-echo tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between teh database and the server, it may also be possibel to use the exceptions tactic between teh server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

@@

Due to the lowing price of hardware and the increasing price of developement time, it is often cheapest and certainly simper to use the tactic \emph{Increase available resourcez}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cach the entire result in preperation for a search refinement. If the database is remotely located the caching of results becomes even more important.

There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under developement control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconcivably expensive compare to simply increasing the computational resources.

\subsection{Quality Attributes - Question 4}

\begin{question}
Discuss where Service-Oriented Architecture can play a role
with respect to the Beer Web Store? Consider the quality implications
\end{question}

As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sens is naturally between teh components.

Instead of creating a web-server which access a database through a specific facade, one may argue whether the desired functionality may not be generalised. "Look-up in database". This may easily be hidden behind a standardised service, e.g. a web-service, meaning that anyone on the WWW who supplies a batabase service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.

Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service. This will also point the way for a more standardised lookk-and-feel when shopping for beer.

Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advabtages.

Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be benificial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.

Naturally this generality do not necesarrliy come free. A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly. On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby redusing the chance of the system becomming fully non-responsive. Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.

@@

The quality attributes are vital to the software architecture, as they not only define the requirements to the software architecture (along with the scenarios), but also describe their relative importance. The quality attributes therefore allow the architecture the be evaluated in terms of how well it supports or enables a system's desired external (e.g. availability and performance) and internal behaviour (e.g. testability and modifiability). \cite{h_sarjoughian2002}

>>>>

Once the scenarios are clarified and voted on, then the top priority scenarios may be formalized and it may be determined which quality attribute(s) they involve. For examples lets assume that the above scenario was selected as top priority.

As it may be seen, the scenario(s) focus primarily on \emph{Availability} (how rare should it be that a user experience non-responsiveness and how many simultaneous users can the system support), \emph{Usability} (how easy should it be for the user to cancel a transaction halfway through and how easy should it be to pick up a previous transaction), \emph{Performance} (how big a delay can the user expect when performing an operation like a lookup in the database), \emph{Security} (how to avoid someone using stolen credit cards and also prevent a "hacker" from stealing the credit card information either during the transaction or from the Beer Web Store's database.

@@

Finding out if a resource is down may be done using one of the fault detection tactics; ping/echo, heartbeat or exception. Due to the loose coupling between the client (browser) and the server, the exception solution is not beneficial. Since it may be assumed that the supervising process (remotely deployed) (the process monitoring if the server is down) is stable, then a heartbeat is sufficient to ensure availability, meaning that the web-server will issue a heart-beat every e.g. 1 second to a maintenance process, and if one (or perhaps more) heartbeats is missed then the maintenance service can act accordingly.

The proper response to a faulty server may be many things, as defined by the tactics group \emph{fault recovery}. As the requirement for availability stipulated a three stage fault recovery, the server or servers (as it may be required to have multiple serves to do load-balancing, depending on how many users must be supported) can service the 5 second requirement, and then if an error occur (as detected by the heartbeat), a restart or swap to another server/database can be performed in 30 seconds. As 30 seconds is insufficient to boot a web server and a database, the \emph{Spare} tactic would be insufficient. At the same time having \emph{Active redundancy} would be over-kill, as this kind of immediate switch-over is not required. For these reasons \emph{Passive Redundancy} is sufficient.

As it is an Internet system, it will also be sufficient to simply restart the session, and no \emph{Shadow}, \emph{State resynchronization} or \emph{Rollback} is needed in the main web server. The user may experience being thrown back to the main page, yet this is deemed acceptable, under availability and performance, yet it may not be in Usability. During a payment transaction, however, it must be ensured that the purchase is either completed or not-completed - halfway completed is not acceptable. For this it will be beneficial to use the \emph{Rollback} tactic on the \emph{Basket} component.

Finally a web-server may be deployed multiple times on the same machine, allowing for the recovery or prevention of the lesser faults, by simply re-spawning the process that has become non-responsive, utilizing the \emph{Process Monitor} tactic. This is e.g. done by the Apache-server.

As it is not only the web server which may become non-responsive, it is also required to monitor the database. However, as a non-responsive data-base may be discovered by a non-responsive server, it is possible to use the \emph{ping-echo} tactic with the database, and only perform the ping-echo if the server becomes non-responsive, to find out how much must be restarted (only the server or both the server and the database). 
Depending on the binding between the database and the server, it may also be possible to use the exceptions tactic between the server and the database, so the server itself can restart the database, if an exception occur in a call to the database.

@@

Due to the decreasing price of hardware and the increasing price of development time, it is often cheapest and certainly simpler, to use the tactic \emph{Increase available resource}; introduce load balancing with multiple parallel servers thereby increase performance during heavy-load situations, ensure that the serves have sufficient network bandwidth, RAM and CPU strength to handle many simultanious database look-ups - here the tactic of \emph{Maintain multiple copies of data or computations} may be used to speed up database-look-up by caching the most popular searches, or cache the entire result in preparation for a search refinement. If the database is remotely located the caching of results becomes even more important.

There are many other tactics which may be employed, yet given the nature to the application, it is not believed that the biggest resource consumers are under development control, for building the web-server from scratch to support better concurrency or using non-standard performance tweaking algorithms for the database-lookup, will most likely become inconceivably expensive compare to simply increasing the computational resources.

\subsection{Quality Attributes - Question 4}
\label{sec:qa_q4}
\begin{question}
Discuss where Service-Oriented Architecture can play a role
with respect to the Beer Web Store? Consider the quality implications
\end{question}

As SOA is all about standardised interfaces supplying a functionality which may be used by many different products, the areas where SOA makes sense is naturally between the components.

Instead of creating a web-server which access a database through a specific fa\c cade, one may argue whether the desired functionality may not be generalised. "Look-up in database". This may easily be hidden behind a standardised service, e.g. a web-service, meaning that anyone on the WWW who supplies a database service may be used to host the beer database, and if one becomes dissatisfied with the database provider, it is transparent to simply switch to another.

Taking this one step further, perhaps the database service may be extended with a data-base type, meaning that the Beer Web Store's domain model is exposed in a service, which would allow for not only displaying ones own brands of beer, but through a discovery service, also beer-brands by other brewers on the net who implement this service. This will also point the way for a more standardised look-and-feel when shopping for beer.

Naturally just because a web-service (or other service) is used it does not mean that it has to be public, jet it does have some advantages.

Finally the WS-Security web-service may be used to handle payment, and also possibly, if the database or Beer Web Store Domain model is exposed through a web-service, it may also be beneficial that not everyone may have the ability to update the domain model or database, so here the WS-Security service may also come in handy.

Naturally this generality do not necessarily come free. A look-up in a database behind a service, especially a web-service, is far slower than a direct look-up in a local database, thereby influencing Performance poorly. On the other hand the use of web-services may increase Availability, as the functionality is deployed on multiple servers, thereby reducing the chance of the system becoming fully non-responsive. Naturally the critical path must be evaluated, as if a link on the critical path chain become non-responsive then the entire system becomes non-responsive.

@@

The quality attributes are vital to the software architecture, as they not only define the requirements to the software architecture (along with the scenarios), but also describe their relative importance. The quality attributes therefore allow the architecture to be evaluated in terms of how well it supports or enables a system's desired external (e.g. availability and performance) and internal behaviour (e.g. testability and modifiability). \cite{h_sarjoughian2002}
